An open-source integrated framework for the automation of citation collection and screening in systematic reviews.

### 2022-02-22 Manuscript/Manuscript.tex
more language check and updated abstract

\begin{abstract}
The exponential and prolific growth of published original scientific
contributions makes secondary literature abridgements increasingly
demanding. We introduce a new open-source framework for systematic
reviews that significantly reduces time and human resource allocation
for text identification, collection, and screening of scientific
literature.\\
The framework provides three main tools: 1) an automatic citation search
engine and manager that allows collecting records from multiple citation
databases (Pubmed, WOS, SCOPUS, EMBASE, IEEE) with a unified query
syntax, 2) a citation screening tool based on Bayesian active machine
learning and natural language processing that requires users feedback
only on uncertain classifications to iteratively increase predictive
accuracy and, 3) a semi-automatic, data-driven query generator to create
new search queries from existing reviewed citation data sets.\\
The framework was tested on an example topic to collect and screen
citations of related scientific production. To evaluate the performance
of the active machine learning classifier, we estimated the median
posterior sensitivity and efficiency {[}90\% Credible Intervals{]} using
Bayesian simulation to predict the distribution of potentially relevant
matches among the records not manually reviewed.\\
A total of 17,755 unique records were collected through the framework
citation manager; 101 of 766 records were found to be relevant after
manual evaluation, while the rest were excluded by automatic
classification; the theoretical efficiency was 95.6\% {[}95.3\%,
95.7\%{]} with a sensitivity of 100\% {[}93.5\%, 100\%{]}.\\
A new search query was generated from the labelled dataset, and 82,579
additional records were collected; only 567 records required human
review after automatic screening, and six additional positive matches
were found. Including the additional records, the overall expected
sensitivity decreased to 97.3\% {[}73.8\%, 100\%{]} while the efficiency
increased to 98.6\% {[}98.2\%, 98.7\%{]}.\\
For large studies, the framework can significantly reduce the human
resources required to conduct systematic reviews by simplifying citation
collection and screening while demonstrating exceptional sensitivity.
Such a tool can improve the standardization and repeatability of
systematic reviews.
\end{abstract}

@@

by the COVID-19 pandemic, thanks to faster article' processing time by

@@

1999). A consequence of this is systematic reviews are often already
out-of-date by the time they are published (E. M. Beller et al. 2013).\\

@@

In this manuscript, we present an open source, production-ready
framework that further contributes to the state-of-the-art in systematic
review automation (SRA) and helpers (SRH) tools. We improve the
``retrieval phase'' by providing a unified framework for the automated
collection and management of scientific literature from multiple online
sources. For the citation screening (appraisal) phase, we built an
active machine learning-based protocol (Settles 2009; Miwa et al. 2014),
which utilises a Bayesian framework to efficiently identify potentially
relevant documents that require human review, while automatically
screening-out the vast majority of clearly non-relevant ones; the
algorithm then requires human review to iteratively increase
classification accuracy. Finally, we included a tool to generate new
search queries based on an already labelled citation data set, to
identify relevant research that may possibly have been missed by
manually-made queries.\\

@@

We built an R (R Core Team 2020) based framework to simplify two aspects
of systematic literature review: record acquisition and classification.
The code used to generate the results in the manuscript is available at
\url{https://github.com/AD-Papers-Material/BART_SystReviewClassifier},
while an updated and ready to use version of the framework is
distributed as an R package at
\url{https://github.com/bakaburg1/BaySREn}. The framework includes
several modules that communicate through intermediate outputs stored in
standard formats, which make it possible for users to extend the
framework or easily integrate it with other tools in their pipeline. See
Supplemental Material S1 for an in-depth description of the framework
and how to use it.\\
The tasks carried out by the framework are grouped into ``sessions'',
which comprise obtaining scientific citation data (records) using a
search query and then labelling them as relevant (``positive'' in the
rest of the text) or not (``negative'') for the topic of interest with
the help of a machine learning engine (Fig. 1). The initial search query
should be built using domain knowledge, trying to achieve a high
relevant/non-relevant record ratio.\\
The framework can then generate a new data-driven query from this
labelled set to perform a new session to find records possibly missed by
the first query.\\

\begin{figure}
\includegraphics[width=1\linewidth]{methods_diagram} \caption{Figure 1.Framework's visual depiction.}\label{fig:method_diagram}
\end{figure}

@@

We built a set of tools to allow users to automatically search and
download citation data from three major scientific databases
(``sources''): Pubmed (\url{https://pubmed.ncbi.nlm.nih.gov/}), Web Of
Science (WOS, \url{https://apps.webofknowledge.com/}) and the Institute
of Electrical and Electronics Engineers (IEEE,
\url{https://ieeexplore.ieee.org/Xplore/home.jsp}). The framework takes
care of authorization management for non-open databases like WOS and
IEEE. It is also possible to download and import records in the
framework manually; this is particularly useful to acquire records from
the SCOPUS (\url{https://www.scopus.com/}) and EMBASE databases
(\url{https://www.embase.com/}), for which a comprehensive API interface
was not easy to build. An extra manual search was also necessary for
Pubmed, since the API and the web interface have different rule
expansion algorithms and return slightly different results ({`NCBI
Insights : Updated Pubmed e-Utilities Coming in April 2022!'}, n.d.). A
short guide on how to set up the framework for each database supported
is available in Supplemental Material S3.\\
The collected records are merged into a single database, resolving
duplicates and different formatting between sources. The records are
ordered according to the frequency of the positive query terms (e.g.,
not preceded by a \emph{NOT} modifier) in the title and abstract
(``simple query ordering'').\\
The researcher is then asked to label a number of records to create the
``initial training set'' needed to start the automatic classification.
We recommend manually labelling the first 250 records (see
``hyperparameter optimization'' later). Simple query ordering increases
the positivity rate in the initial training set (Wallace, Small, et al.
2010), leading to higher sensitivity during automatic classification
(Chawla, Japkowicz, and Kotcz 2004).

@@

The collected citation data contain a number of fields characterising a
scientific publication. The framework models the relevance of a record
based on the following fields: title, abstract, authors, keywords, MESH
terms (Lipscomb 2000). A range of Natural Language Processing (NLP)
techniques (Baeza-Yates, Ribeiro-Neto, et al. 1999; Marshall and Wallace
2019; Ananiadou and McNaught 2006) are employed to convert the textual
information in these fields into features for machine learning through a
bag-of-words approach (Marshall and Wallace 2019). Processing of free
text fields (title, abstract) includes: tokenization (i.e., extracting
the terms), removal of common stopwords (i.e., sentence components
having no semantic value), part-of-speech filtering (only nouns,
adjectives, verbs and untagged terms are retained), and lemmatisation of
terms (i.e., reduction to their base grammatical form). Text processing
for authors, keywords and MESH terms identifies logical units (e.g.,
authors' full names, composite keywords) and extracts them.\\
Terms appearing in less than 5\% of the labelled documents are removed
from negative records. All terms in the positive set are kept to
increase sensitivity at the cost of specificity.\\
Some terms tend to co-appear in records (non-consecutive ngrams,
nc-ngrams), often carrying a particular meaning when they do co-occur.
To detect nc-ngrams, we generated a word network representation
(Francois Rousseau 2015) with edges occurring between terms with a
cosine similarity in terms of docuement co-occurrence \textgreater{}
0.5. We extracted the maximal cliques in the network (Eppstein, Löffler,
and Strash 2010) representing highly correlated groups of terms; these
groups are added to the dataset as individual features. Only nc-ngrams
comprising a maximum of ten terms are kept.\\
A second network is built using a co-occurrence threshold of 0.9. In
this case, the cliques represent terms that always appear together and
can therefore be considered redundant (i.e., they do not need to be
considered separately). These terms are merged to increase computational
efficiency and reduce overfitting.\\
The output is a Document-Term Matrix (DTM), with \(N_d\) rows
representing the records (\(D_i\)), \(N_t\) terms column for the
\(t_{field}\) terms (divided by record field) and \({0,1}\) values
whether \(t_{field} \in D_i\). We also enriched the DTM with features
referencing the number of terms in each field to help the model scale
term importance based on the field length.

@@

We used a Bayesian Additive Regression Trees (BART) machine learning
``classification model'' (Chipman et al. 2010) (in the implementation of
Kapelner and Bleich 2013) to predict the probability of a record being
relevant, given the information coded into the enriched DTM and the
initial training set. We set up the BART model to use 2,000 MCMC
iterations (after 250 burn-in iterations) and 50 trees; we used a \(k\)
value of 2 to regularize extreme prediction and let the model use
missing fields in the DTM as features (Kapelner and Bleich 2015).
Positive records are oversampled ten times to increase sensitivity
(Batista, Prati, and Monard 2004).\\
The output is a posterior predictive distribution (PPD) for each record
describing the probability of it being relevant (i.e., a positive
match). An ensemble of ten models was fitted to improve prediction
stability by averaging the PPD between models (Zhou 2021; Dietterich
2000).\\

To assign the labels, we employed an ``active learning'' approach
(Settles 2009; Miwa et al. 2014), where a human reviews a specific
subset of predictions made by the machine, which is then retrained on
the manually reviewed dataset. This process is carried out iteratively
to reduce prediction uncertainty.\\
Label assignment is done through identification of an ``uncertainty
zone'' the construction of which is possible thanks to the Bayesian
nature of BART, which provides full PPDs instead of point-wise
predictions for each record.\\
To describe the process formally, we define

@@

as the PPD of a record \(D_i\) being relevant (i.e, having a positive
label, \(L_i = 1\)), averaging the PPDs of the ensemble of \(M=10\)
models \(m\), and

@@

respectively as the lower and upper boundaries of the 98\% quantile
interval of \(\pi_i\) (98\% Predictive Interval, 98\% PrI).\\
Then we identify the ``uncertainty zone'' as

\[U_\pi=[\max\vec{\pi}_{u}^-, \min\vec{\pi}_{l}^+]\] with
\(\vec{\pi}_{u}^-\) being the vector of \(\pi_{i,u}\) with a negative
label and \(\vec{\pi}_{l}^+\) the vector of \(\pi_{i,l}\) with a
positive label. That is, \(U_\pi\) defines a range of values between the
smallest \(\pi_{i,l}\) in the set of already labelled positive records
\(L_p\) and the largest \(pi_{i,u}\) related to the negative ones
\(L_n\), noting that the two limits can appear in any order.\\
Consequently, a record \(D_i\) will be labelled as positive if

@@

that is, the record lower 98\% PrI boundary should be higher than every
value in the uncertainty zone. In other words, for a record to be
labelled positive, its PPD should be within the range of the mixture of
PPD of the records previously labelled positive, and should not cross
the distributions of the negative records.\\
Conversely, a record is labelled as negative if

@@

With the BART model, it is possible to extract the proportion of times a
term is used in a posterior tree over the sum of total inclusions of all
variables (Kapelner and Bleich 2013), as a measure of importance of the
features in the model. We extracted the terms, the portion of the
citation data in which they were used, the average ``inclusion rate''
among the ensemble models (over 10,000 inclusions) and its ratio over
the standard deviation of this inclusion (inclusion stability, IS). For
each term we ran a Poisson regression to get the linear association with
a positive label and reported it as Relative Risk (RR) with the number
of standard errors as significance index (Statistic); the comparison
between the inclusion rate in the BART models and the linear association
allows to spot relevant non-linear effects (i.e, the feature is relevant
only in association with others). In the Results, we only listed the
first 15 terms with IS \textgreater{} 1.5, (in order of inclusion rate),
while the first fifty terms regardless of inclusion stability are listed
in Supplemental Material S2.

@@

We developed an algorithm that generates a new search query to find
further relevant publications missed in the initial search, possibly at
a reasonable cost to specificity (i.e., a higher number of negative
results).\\
The algorithm encompasses a number of steps:

\begin{itemize}
\tightlist
\item
  A partition tree (Therneau and Atkinson 2019) is built between the DTM
  and 800 samples from the PPD; if a term is present multiple times in
  the DTM (e.g., both in the title and abstract), it is counted just
  once, and field term count features are removed. This step generates a
  list of rules composed by \emph{AND}/\emph{NOT} ``conditions'' made of
  terms/authors/keywords/MESH tokens, which together identify a group of
  records.
\item
  For each rule, negative conditions (i.e., \emph{NOT} statements) are
  added iteratively, starting from the most specific one, until no
  conditions are found that would not also remove positive records.
\item
  The extended set of rules is sorted by positive-negative record
  difference in descending order. The cumulative number of unique
  positive records is computed and used to group the rules. Rules inside
  each group are ordered by specificity.
\item
  The researcher is then asked to review the rule groups and select one
  or more rules from each group (useful if they convey different
  meaning), or edit them if overly too specific positive or negative
  conditions were included. It is possible to exclude a group of rules
  altogether, especially those with the poorest sensitivity/specificity
  ratio.
\item
  The selected rules are joined together by \emph{OR} statements,
  defining a subset of records with a sensibly higher proportion of
  positive records than the original set
\item
  Redundant (i.e., rules whose positive records are already included in
  more specific ones) and non-relevant rules (i.e., conditions that when
  removed do not impact sensitivity and specificity) are removed.
\item
  Finally, the rules are re-elaborated in a query that can be used in
  the major scientific databases.
\end{itemize}

Because the algorithm is data-driven, it creates queries that
effectively select positive records from the input dataset, but may be
not specific enough when applied to actual research databases. Therefore
we added an extra subquery in \emph{AND}, that specifies the general
topics of our search and narrows the search domain.\\
The new query was used to initiate a second search session.

@@

We trained a simple Bayesian logistic regression (surrogate model) on
the reviewed records to evaluate the consistency of classification model
(see Discussion for the theoretical justification). The surrogate model
uses as predictor the lower boundary of the 98\% PrI of the PPD of the
records with weakly regularizing, robust priors for the intercept
(Student T with \(\nu=3,\mu=0,\sigma=2.5\)) and the linear coefficient
(Student T with \(\nu=3,\mu=0,\sigma=1.5\)).\\
The quality of the model was evaluated through the Bayesian \(R^2\)
(Gelman et al. 2019), of which we reported the posterior median and 90\%
Credible Interval {[}90\% CrI{]}. The \(R^2\) also provides an
evaluation of the consistency of the original classification model.
Given that this model is conditional only on the BART predictions and
not on the DTM, it is characterized by more uncertainty, providing
plausible worst-case scenarios.\\
The surrogate model is then used to generate the predictive cumulative
distribution of the number of total positive records in the whole
dataset. This distribution allows to estimate the expected total
posterior ``Sensitivity'' and ``Efficiency'' of the classification model
in the full (unreviewed) dataset. Efficiency is summarised by the ``Work
saved over random'' (WSoR) statistic: one minus the ratio between the
number of record manually reviewed and those that would be required to
find the same number of positives if classification were performed
choosing records randomly; this last quantity is estimated through a
negative hypergeometric distribution (Chae 1993) over the predicted
number of positive records.\\
For the number of predicted positive records, sensitivity and
efficiency, we reported the ``truncated 90\% PrI'' {[}trunc. 90\%
PrI{]}, i.e., the uncertainty interval bounded by the number of observed
total positive records (i.e., there cannot be fewer predicted positive
records than observed).

@@

Our classification algorithm has a number of hyperparameters:

@@

To evaluate the hyperparameter effect of performance, we set up a ``grid
search'' (Claesen and De Moor 2015; L. Yang and Shami 2020) on a
prelabelled ``validation set'' derived from the first 1,200 records of
the first session dataset. Each hyperparameter combination was tested
until four CR iterations were completed with no positive records or the
whole dataset was labelled.\\
For each combination, a performance score was computed as the product of
``Efficiency'' (1 minus the ratio of records that required reviewing
over the total number of records) and ``Sensitivity'' (number of
positive records found over the total number of positive records). We
then used a partition tree (Therneau and Atkinson 2019) to identify
homogeneous ``performance clusters'' of scores given hyperparameter
values. For the rest of the study, we used the best hyperparameter set
in terms of Sensitivity followed by Efficiency from the cluster with the
highest average score.\\

@@

The first search session returned a total of 27,600 records,
specifically 12,719 (71.6\% of the total) records from the EMBASE
database, followed by 9,546 (53.8\%) from Pubmed, 3,175 (17.9\%) from
SCOPUS, 2,100 (11.8\%) from WOS, and 60 (0.34\%) from IEEE (Table 1).
There were various degrees of overlapping between sources, with 38.4\%
of records being present in more than one database, and EMBASE and IEEE
being the databases with the higher uniqueness ratios. The final data
set was composed from 17,755 unique records.\\
The first 250 records (based on ``simple query ordering'') were labeled
manually. Of these 43 (17.2\%) were labeled as positive, and 207
(82.8\%) as negative.

@@

The ``\&'' indicates nc-ngrams, i.e., terms strongly co-occurrent in the
documents.\\
The engine was able to pick up the central concept of the research
topic, i.e., ``patient transport'' or ``transfer'' through a ``network''
of ``facility''ies that facilitates the ``spread'' of infections, and
even one of the authors of this study (Donker T.) as well as the region
of interest (``Orange County'') of another research group active on the
topic of pathogen spreading over hospital networks. Some terms were
considered highly relevant by the BART models (e.g., ``Worker'' in 6th
position out of more than 3800 terms considered) although in a simpler
linear model their effect would hardly be significant (statistic: -1.21
s.e.); these are terms that are only relevant in conjunction with other
terms but not by their own, highlighting the extra predictive power
achievd through the use of advanced, non-linear machine learning.\\
A more extensive set of terms is presented in Table 1 of the
Supplemental Material S2.

@@

The results of the first classification session were used to create a
second, data-driven query with the objective of performing a more
extensive search to find records that may have been missed during the
first search session. The resulting query was as follows:\\

@@

The final piece \emph{AND ((antimicrobial resistance) OR (healthcare
infection)} was added manually to better define the search domain, since
the algorithm was trained on documents that were all more or less
related to these topics.\\
The generated query also provides a more nuanced understanding of the
engine's internal classification logic, and this is helpful to spot
possible biases in the model.\\

@@

The new search produced 107,294 records (Table 1), of which 48,396
(58.6\%) from the EMBASE, followed by 28,811 (34.9\%) from Pubmed,
17,070 (20.7\%) from SCOPUS, 12,956 (15.7\%) from WOS, and 61 (0.074\%)
from IEEE; compared with the first session, the relative weight of
EMBASE and Pubmed decreased, while the level of content specificity
greatly increased, as it was for SCOPUS. After removal of duplicates,
82,579 unique records were obtained. The newly collected records were
joined with those from the first session and duplicates were removed. We
obtained 98,371 unique records, with just 1,963 shared records between
searches, which equates to 2\% of the total. The percentage of records
shared by two or more source dropped to 22\%.

>>>>

\begin{abstract}
The exponential growth of scientific production makes secondary
literature abridgements increasingly demanding. We introduce a new
open-source framework for systematic reviews that significantly reduces
time and workload for collecting and screening scientific literature.\\
The framework provides three main tools: 1) an automatic citation search
engine and manager that collects records from multiple online sources
with a unified query syntax, 2) a Bayesian, active machine learning,
citation screening tool based on iterative human-machine interaction to
increase predictive accuracy and, 3) a semi-automatic, data-driven query
generator to create new search queries from existing citation data
sets.\\
To evaluate the automatic screener's performance, we estimated the
median posterior sensitivity and efficiency {[}90\% Credible
Intervals{]} using Bayesian simulation to predict the distribution of
undetected potentially relevant records.\\
Tested on an example topic, the framework collected 17,755 unique
records through the citation manager; 766 records required human
evaluation while the rest were excluded by the automatic classifier; the
theoretical efficiency was 95.6\% {[}95.3\%, 95.7\%{]} with a
sensitivity of 100\% {[}93.5\%, 100\%{]}.\\
A new search query was generated from the labelled dataset, and 82,579
additional records were collected; only 567 records required human
review after automatic screening, and six additional positive matches
were found. The overall expected sensitivity decreased to 97.3\%
{[}73.8\%, 100\%{]} while the efficiency increased to 98.6\% {[}98.2\%,
98.7\%{]}.\\
The framework can significantly reduce the workload required to conduct
large literature reviews by simplifying citation collection and
screening while demonstrating exceptional sensitivity. Such a tool can
improve the standardization and repeatability of systematic reviews.
\end{abstract}

@@

by the COVID-19 pandemic, thanks to faster article processing time by

@@

1999). A consequence of this is that systematic reviews are often
already out-of-date by the time they are published (E. M. Beller et al.
2013).\\

@@

This manuscript introduces an open-source, production-ready framework
that further contributes to the state-of-the-art in systematic review
automation (SRA) and helpers (SRH) tools. We improve the ``retrieval
phase'' by providing a unified framework for the automated collection
and management of scientific literature from multiple online sources.
For the citation screening (appraisal) phase, we built an active machine
learning-based protocol (Settles 2009; Miwa et al. 2014), which utilises
a Bayesian framework to efficiently identify potentially relevant
documents that require human review while automatically screening-out
the vast majority of clearly non-relevant ones; the algorithm then
requires human review to increase classification accuracy iteratively.
Finally, we included a tool to generate new search queries based on an
already categorised citation data set, to identify relevant research
that manually-made queries may have possibly missed.\\

@@

We built an R (R Core Team 2020) based framework to simplify two aspects
of systematic literature review: record acquisition and classification.
The code used to generate the results is available at
\url{https://github.com/AD-Papers-Material/BART_SystReviewClassifier},
while an updated and ready to use version of the framework is
distributed as an R package at
\url{https://github.com/bakaburg1/BaySREn}. The framework includes
several modules that communicate through intermediate outputs stored in
standard formats, making it possible for users to extend the framework
or easily integrate it with other tools in their pipeline. See
Supplemental Material S1 for an in-depth description of the framework
and how to use it.\\
The tasks carried out by the framework are grouped into ``sessions'',
which comprise obtaining scientific citation data (records) using a
search query and then labelling them as relevant (``positive'' in the
rest of the text) or not (``negative'') for the topic of interest with
the help of a machine learning engine (Fig. 1). The initial search query
should be built using domain knowledge, trying to achieve a high
relevant/non-relevant record ratio.\\
The framework can then generate a new data-driven query from this
labelled set to perform a new session to find records possibly missed by
the first query.\\

\begin{figure}
\includegraphics[width=1\linewidth]{methods_diagram} \caption{Figure 1. Framework's visual representation.}\label{fig:method_diagram}
\end{figure}

@@

We built a set of tools to allow users to automatically search and
download citation data from three major scientific databases
(``sources''): Pubmed (\url{https://pubmed.ncbi.nlm.nih.gov/}), Web Of
Science (WOS, \url{https://apps.webofknowledge.com/}) and the Institute
of Electrical and Electronics Engineers (IEEE,
\url{https://ieeexplore.ieee.org/Xplore/home.jsp}). The framework
handles authorisation management for non-open databases like WOS and
IEEE. It is also possible to import previously downloaded records in the
framework; this is particularly useful for acquiring records from SCOPUS
(\url{https://www.scopus.com/}) and EMBASE databases
(\url{https://www.embase.com/}), for which a comprehensive API interface
was not easy to build. An extra manual search was also necessary for
Pubmed since the API and the web interface have different rule expansion
algorithms and return slightly different results ({`NCBI Insights :
Updated Pubmed e-Utilities Coming in April 2022!'}, n.d.). A short guide
on how to set up the framework for each database supported is available
in Supplemental Material S3.\\
The collected records are merged into a single database, resolving
duplicates and different formatting between sources. The records are
ordered according to the frequency of the positive query terms (e.g.,
not preceded by a \emph{NOT} modifier) in the title and abstract
(``simple query ordering'').\\
The researcher is then asked to label a subset of records to create the
``initial training set'' needed to start the automatic classification.
We recommend manually labelling the first 250 records (see
``hyperparameter optimisation'' later). Simple query ordering increases
the positivity rate in the initial training set (Wallace, Small, et al.
2010), leading to higher sensitivity during automatic classification
(Chawla, Japkowicz, and Kotcz 2004).

@@

The framework models the relevance of a record based on the following
fields in the citation data: title, abstract, authors, keywords, MESH
terms (Lipscomb 2000). A range of Natural Language Processing (NLP)
techniques (Baeza-Yates, Ribeiro-Neto, et al. 1999; Marshall and Wallace
2019; Ananiadou and McNaught 2006) are employed to convert the textual
information in these fields into features for machine learning through a
bag-of-words approach (Marshall and Wallace 2019). Processing of free
text fields (title, abstract) includes: tokenisation (i.e., extracting
the terms), removal of common stopwords (i.e., sentence components
having no semantic value), part-of-speech filtering (only nouns,
adjectives, verbs and untagged terms are retained), and lemmatisation of
terms (i.e., reduction to their base grammatical form). Text processing
for authors, keywords and MESH terms identifies logical units (e.g.,
authors' full names, composite keywords) and extracts them.\\
Terms appearing in less than 5\% of the labelled documents are removed
from negative records. All terms in the positive set are kept to
increase sensitivity at the cost of specificity.\\
Some terms tend to co-appear in records (non-consecutive ngrams,
nc-ngrams), often carrying a particular meaning when they do co-occur.
To detect nc-ngrams, we generated a word network representation
(Francois Rousseau 2015) with edges occurring between terms with a
cosine similarity in terms of document co-occurrence \textgreater{} 0.5.
We extracted the maximal cliques in the network (Eppstein, Löffler, and
Strash 2010) representing highly correlated groups of terms; these
groups are added to the dataset as individual features. Only nc-ngrams
comprising a maximum of ten terms are kept.\\
A second network is built using a co-occurrence threshold of 0.9. In
this case, the cliques represent terms that always appear together and
can therefore be considered redundant (i.e., they do not need to be
considered separately). These terms are merged to increase computational
efficiency and reduce overfitting.\\
The output is a Document-Term Matrix (DTM), with \(N_d\) rows
representing the records (\(D_i\)), \(N_t\) terms column for the
\(t_{field}\) terms (divided by record field) and \({0,1}\) values
whether \(t_{field} \in D_i\). We also enriched the DTM with features
referencing the number of terms in each field to help the model scale
term importance based on the field length.

@@

We used a Bayesian Additive Regression Trees (BART) machine learning
``classification model'' (Chipman et al. 2010) (in the implementation of
Kapelner and Bleich 2013) to predict the probability of a record being
relevant, given the information coded into the enriched DTM and the
initial training set. We set up the BART model to use 2,000 MCMC
iterations (after 250 burn-in iterations) and 50 trees; we used a \(k\)
value of 2 to regularise extreme prediction and let the model use
missing fields in the DTM as features (Kapelner and Bleich 2015).
Positive records are oversampled ten times to increase sensitivity
(Batista, Prati, and Monard 2004).\\
The output is a posterior predictive distribution (PPD) for each record
describing the probability of it being relevant (i.e., a positive
match). An ensemble of ten models was fitted to improve prediction
stability by averaging the PPD between models (Zhou 2021; Dietterich
2000).\\

To assign the labels, we employed an ``active learning'' approach
(Settles 2009; Miwa et al. 2014), where a human reviews a specific
subset of predictions made by the machine, which is then retrained on
the manually reviewed dataset. This process is carried out iteratively
to reduce prediction uncertainty.\\
Label assignment is done through identification of an ``uncertainty
zone'', the construction of which is possible thanks to the Bayesian
nature of BART, which provides full PPDs instead of point-wise
predictions for each record.\\
To describe the process formally, we define:

@@

as the PPD of a record \(D_i\) being relevant (i.e, having a positive
label, \(L_i = 1\)), averaging the PPDs of the ensemble of \(M=10\)
models \(m\), and:

@@

respectively as the lower and upper boundaries of the 98\% quantile
interval of \(\pi_i\) (98\% Predictive Interval, 98\% PrI).\\
Then we identify the ``uncertainty zone'' as:

\[U_\pi=[\max\vec{\pi}_{u}^-, \min\vec{\pi}_{l}^+]\]

with \(\vec{\pi}_{u}^-\) being the vector of \(\pi_{i,u}\) with a
negative label and \(\vec{\pi}_{l}^+\) the vector of \(\pi_{i,l}\) with
a positive label. That is, \(U_\pi\) defines a range of values between
the smallest \(\pi_{i,l}\) in the set of already labelled positive
records \(L_p\) and the largest \(pi_{i,u}\) related to the negative
ones \(L_n\), noting that the two limits can appear in any order.\\
Consequently, a record \(D_i\) will be labelled as positive if:

@@

that is, the record lower 98\% PrI boundary should be higher than every
value in the uncertainty zone. In other words, for a record to be
labelled positive, its PPD should be within the range of the mixture of
PPD of the records previously labelled positive and should not cross the
distributions of the negative records.\\
Conversely, a record is labelled as negative if:

@@

As a measure of feature importance, we computed the ``inclusion rate'',
that is, the proportion of times a term is used in a posterior tree over
the sum of total inclusions of all variables (Kapelner and Bleich 2013).
We extracted the terms, the portion of the citation data in which they
were used, the average ``inclusion rate'' among the ensemble models
(over 10,000 inclusions) and its ratio over the standard deviation of
this inclusion (inclusion stability, IS). For each term, we ran a
Poisson regression to get the linear association with a positive label
and reported it as Relative Risk (RR) with the number of standard errors
as significance index (Statistic); the comparison between the inclusion
rate in the BART models and the linear association allows to spot
relevant non-linear effects (i.e., the feature is relevant only in
association with others). In the Results, we only listed the first 15
terms with IS \textgreater{} 1.5 (in order of inclusion rate), while the
first fifty terms, regardless of inclusion stability, are listed in
Supplemental Material S2.

@@

We developed an algorithm that generates a new search query to find
further relevant publications missed in the initial search, possibly at
a reasonable cost to specificity (i.e., a higher number of negative
results).\\
The algorithm is composed of the following steps:

\begin{itemize}
\tightlist
\item
  A partition tree (Therneau and Atkinson 2019) is built between the DTM
  and 800 samples from the PPD; if a term is present multiple times in
  the DTM (e.g., both in the title and abstract), it is counted just
  once, and field term count features are removed. This step generates a
  list of rules composed by \emph{AND}/\emph{NOT} ``conditions'' made of
  terms/authors/keywords/MESH tokens, which together identify a group of
  records.
\item
  For each rule, negative conditions (i.e., \emph{NOT} statements) are
  added iteratively, starting from the most specific one, until no
  conditions are found that would not also remove positive records.
\item
  The extended set of rules is sorted by positive-negative record
  difference in descending order. The cumulative number of unique
  positive records is computed and used to group the rules. Rules inside
  each group are ordered by specificity.
\item
  The researcher is then asked to review the rule groups and select one
  or more rules from each group or edit overly specific rules (e.g.,
  citing a non-relevant concept casually associated with a paper, like a
  numeric value or indicator). It is possible to exclude a group of
  rules altogether, especially those with the poorest
  sensitivity/specificity ratio.
\item
  The selected rules are joined together by \emph{OR} statements,
  defining a subset of records with a sensibly higher proportion of
  positive records than the original set
\item
  Redundant (i.e., rules whose positive records are already included in
  more specific ones) and non-relevant rules (i.e., conditions that when
  removed do not impact sensitivity and specificity) are removed.
\item
  Finally, the rules are re-elaborated in a query that can be used to
  perform a new citation search.
\end{itemize}

Because the algorithm is data-driven, it creates queries that
effectively select positive records from the input dataset but may be
not specific enough when applied to actual research databases. Therefore
we added an extra subquery in \_AND\_that specifies the general topics
of our search and narrows the search domain.\\
The new query was used to initiate a second search session.

@@

We trained a simple Bayesian logistic regression (surrogate model) on
the reviewed records to evaluate the consistency of the classification
model (see Discussion for the theoretical justification). The surrogate
model uses as predictor the lower boundary of the 98\% PrI of the PPD of
the records with weakly regularising, robust priors for the intercept
(Student T with \(\nu=3,\mu=0,\sigma=2.5\)) and the linear coefficient
(Student T with \(\nu=3,\mu=0,\sigma=1.5\)).\\
The quality of the model was evaluated through the Bayesian \(R^2\)
(Gelman et al. 2019), of which we reported the posterior median and 90\%
Credible Interval {[}90\% CrI{]}. The \(R^2\) also provides an
evaluation of the consistency of the original classification model.
Given that this model is conditional only on the BART predictions and
not on the DTM, it is characterised by more uncertainty, providing
plausible worst-case scenarios.\\
The surrogate model is then used to generate the predictive cumulative
distribution of the number of total positive records in the whole
dataset. This distribution allows estimating the expected total
posterior ``Sensitivity'' and ``Efficiency'' of the classification model
in the whole (unreviewed) dataset. Efficiency is summarised by the
``Work saved over random'' (WSoR) statistic: one minus the ratio between
the number of records manually reviewed and those that would be required
to find the same number of positives if classification were performed
choosing records randomly; this last quantity is estimated through a
negative hypergeometric distribution (Chae 1993) over the predicted
number of positive records.\\
For the number of predicted positive records, sensitivity and
efficiency, we reported the ``truncated 90\% PrI'' {[}trunc. 90\%
PrI{]}, i.e., the uncertainty interval bounded by the number of observed
total positive records (i.e., there cannot be fewer predicted positive
records than observed).

@@

Our classification algorithm has a limited number of hyperparameters:

@@

To evaluate the hyperparameter effect of performance, we set up a ``grid
search'' (Claesen and De Moor 2015; L. Yang and Shami 2020) on a
prelabelled ``validation set'' derived from the first 1,200 records of
the first session dataset. Each hyperparameter combination was tested
until four CR iterations were completed with no positive records or
until the whole dataset was labelled.\\
For each combination, a performance score was computed as the product of
``Efficiency'' (1 minus the ratio of records that required reviewing
over the total number of records) and ``Sensitivity'' (number of
positive records found over the total number of positive records). We
then used a partition tree (Therneau and Atkinson 2019) to identify
homogeneous ``performance clusters'' of scores given hyperparameter
values. For the rest of the study, we used the best hyperparameter set
in terms of sensitivity followed by efficiency from the cluster with the
highest average score.\\

@@

The first search session returned a total of 27,600 records,
specifically 12,719 (71.6\% of the total) records from the EMBASE
database, followed by 9,546 (53.8\%) from Pubmed, 3,175 (17.9\%) from
SCOPUS, 2,100 (11.8\%) from WOS, and 60 (0.34\%) from IEEE (Table 1).
There were various degrees of overlapping between sources, with 38.4\%
of records being present in more than one database, and EMBASE and IEEE
being the databases with the higher uniqueness ratios. The final data
set was composed of 17,755 unique records.\\
The first 250 records (based on ``simple query ordering'') were
categorised manually. Of these 43 (17.2\%) were labeled as positive, and
207 (82.8\%) as negative.

@@

The ``\&'' indicates nc-ngrams, i.e., terms strongly co-occurrent in the
documents.\\
The engine was able to pick up the central concept of the research
topic, i.e., ``patient transport'' or ``transfer'' through a ``network''
of ``facility''ies that facilitates the ``spread'' of infections, and
even one of the authors of this study (Donker T.) as well as the region
of interest (``Orange County'') of another research group active on the
topic of pathogen spreading over hospital networks. Some terms were
considered highly relevant by the BART models (e.g., ``Worker'' in the
sixth position out of more than 3800 terms considered), although in a
simple linear model, their effect would hardly be significant
(statistic: -1.21 s.e.); these are terms that are only relevant in
conjunction with other terms but not on their own, highlighting the
extra predictive power achieved through the use of advanced, non-linear
machine learning.\\
A more extensive set of terms is presented in Table 1 of Supplemental
Material S2.

@@

The results of the first classification session were used to create a
second, data-driven query to perform a more extensive search to find
records that may have been missed during the first search session. The
resulting query was as follows:\\

@@

The final piece \emph{AND ((antimicrobial resistance) OR (healthcare
infection)} was added manually to define the search domain better since
the algorithm was trained on documents that were all more or less
related to these topics.\\
The generated query also provides a more nuanced understanding of the
engine's internal classification logic, and this is helpful to spot
possible biases in the model.\\

@@

The new search produced 107,294 records (Table 1), of which 48,396
(58.6\%) from the EMBASE, followed by 28,811 (34.9\%) from Pubmed,
17,070 (20.7\%) from SCOPUS, 12,956 (15.7\%) from WOS, and 61 (0.074\%)
from IEEE; compared with the first session, the relative weight of
EMBASE and Pubmed decreased, while the level of content specificity
greatly increased, as it was for SCOPUS. After removal of duplicates,
82,579 unique records were obtained. The newly collected records were
joined with those from the first session and duplicates were removed. We
obtained 98,371 unique records, with just 1,963 shared records between
searches, which equates to 2\% of the total. The percentage of records
shared by two or more sources dropped to 22\%.
